	\documentclass[11pt]{article}
\usepackage{../cs170}
\usepackage{tabu}
\usepackage{tikz}
\usepackage{algorithm} % Boxes/formatting around algorithms
\usepackage[noend]{algpseudocode} % Algorithms
\usepackage{amsmath}

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=blue,
}

\urlstyle{same} %same
\newcommand{\answer}[1]{{\color{blue} \smallskip\textbf{Solution:} #1}}


\def\title{Homework 2}

\begin{document}
\maketitle

\paragraph*{Instructions:}

You are welcome to form small groups (up to 4 people total) to work through the homework, but you {\bf must} write up all solutions by yourself. List your study partners for homework on the first page, or ``none'' if you had no partners.

If using LaTeX (which we recommend), you may use the homework template linked on this \href{https://piazza.com/class/j5vqc3j229b6u7?cid=21}{Piazza post} to get started. 

Begin each problem on a new page. Clearly label where each problem and subproblem begin. The problems must be submitted in order (all of P1 must be before P2, etc). For questions asking you to give an algorithm, respond in what we will refer to as the \textit{four-part format} for algorithms: main idea, pseudocode, proof of correctness, and running time analysis.

Read the \href{https://piazza.com/class/j5vqc3j229b6u7?cid=72}{Homework FAQ Piazza post} on Piazza before doing the homework for more explanation on the four-part format and other clarifications for our homework expectations.

No late homeworks will be accepted. No exceptions. This is not out of a desire to be harsh, but rather out of fairness to all students in this large course. Out of a total of approximately 12 homework assignments, the lowest two scores will be dropped.

\paragraph{Special Questions:}
\begin{itemize}
\item {\em Shortcut questions}: Short questions are usually easy questions that give you opportunities to practice basic materials. However, we understand that some of you are very familiar with the topics and do not want to spend too much time on easy questions. Therefore, we design shortcut questions for this purpose. A shortcut question usually has multiple parts that build upon each other and are ordered by their difficulty level. You can work on those in order or start from wherever you like. However you only need to submit the last part you are able to solve. For example, if a question has 5 parts (a, b, c, d, e), you are confident about part e, you should submit part e without any of the previous four parts. If you are confident about d but not sure about e, you should submit d for grading purposes. Please clearly indicate in your submission which part you are submitting.
\item {\em Redemption questions}: It is important that you carefully read the posted solutions, even for problems you got right. To encourage this, you have the option of submitting a redemption file, a few paragraphs in which you explain, for each problem you choose to cover, what you did wrong and what the right idea was in your own words (not cutting and pasting from the solution!), and appending it to your homework. For example, suppose that as you review your solutions to HW1, you realize you had misunderstood question 3 and answered it incorrectly. You would write down what you just learned, and then submit it in your HW2 assignment the following week. Because these are mainly for your benefit, feel free to format them however is most useful for you.
\item {\em Extra credit questions}: We might have some extra credit questions in the homework for people who really enjoy the materials. However, please note that you should do the extra credit problems only if you really enjoy working on these problems and want an extra challenge. It is likely not the most efficient manner in which to maximize your score.
\end{itemize}

\newpage
\section*{Due Wednesday, September 13, at 4:59pm} 

\input{hw02body.tex}

\end{document}

